---
title: "p8105_hw1_SC6036"
output: github_document
date: "2026-09-25"
---
## Loading Library
```{r loading library}
library(palmerpenguins)
library(tidyverse)
```

## Loading Data
```{r loading data}
data("penguins", package = "palmerpenguins")
penguins_df <- janitor::clean_names(penguins)
penguins_df
```

# Question 1
```{r Q1}
skimr::skim(penguins_df)
summary(penguins_df)

ncol(penguins_df)
nrow(penguins_df)

mean(penguins_df$flipper_length_mm, na.rm = T)

ggplot(penguins_df, 
       aes(x = bill_length_mm, 
           y = flipper_length_mm, 
           color = species)) + 
  geom_point()
ggsave("scatterplot.png")
```
This data set has 3 factor variables and 3 numeric variables. A few important variable include `island`, `species`, `bill_length_mm`, `bill_depth_mm`, `body_mass_g`, `sex`, and `year`. There are mostly Adelie species (152 total) and mostly from the island Biscoe (168 total). The penguins data set has 344 rows and 8 columns. The mean flipper length, after removing NA, is 200.9.


# Question 2
```{r Q2}
set.seed(1)
# setting seed for reproducibility and creating data frame

example_df = tibble(
  vec_numeric = rnorm(10),
  vec_logical = vec_numeric > 0,
  vec_char = sample(words, 10),
  vec_factor = factor(sample(c("level 1", "level 2", "level 3"), 10, replace = T))
)
example_df

# taking means of variables
example_df |> pull(vec_numeric) |> mean()
example_df |> pull(vec_logical) |> mean()
example_df |> pull(vec_char)    |> mean()
example_df |> pull(vec_factor)  |> mean()
```

`vec_numeric` and `vec_logical` work with the mean function, where the first pulls the average and the second pulls the proportion of `True`. `vec_char` does not work, as it is neither numeric nor logical, therefore returning NA. Lastly, `vec_factor` does not work because a factor is neither numeric nor logical, returning NA. 

```{r, eval = F}
as.numeric(example_df$vec_logical)
as.numeric(example_df$vec_char)
as.numeric(example_df$vec_factor)

example_df |> pull(vec_logical) |> mean()
example_df |> pull(vec_char)    |> mean()
example_df |> pull(vec_factor)  |> mean()
```

`vec_logical` coerces to 0 and 1 which represents T and F, so we are able to take the mean now, returning the mean. `vec_char` does not work because the values have now become NA since they were previously words and cannot become numerical, therefore returning NA. Lastly, `vec_factor` looks like it becomes numerical because the levels are given a number label, but taking the mean returns NA.

This does show how character to numeric coercion does not work, and that a factor is neither numeric nor logical so the taking the mean won't work here either. 
